Highly Available Transactions: Virtues and Limitations

### 2013-04-01 writeup/hat-analysis.tex
small bump
fixed
not british or assuie

\textbf{Read Committed} isolation is particularly important in
practice as it is the default of many DBMSs. Implementations differ,
with some based on long-duration exclusive locks and short-duration
read locks~\cite{gray-isolation} and others based on multiple
versions. The implementations typically have recency and monotonicity
properties beyond the simple meaning of the name, which is what is
expressed in the implementation-agnostic definition: under Read
Committed, transactions should do not access uncommitted or
intermediate versions of data items. This prohibits both ``Dirty
Writes'', as above, and also ``Dirty Reads'' phenomena.  This
isolation is Adya's PL-2, and formalised by prohibiting Adya's $G1a-c$
(or ANSI's $P1$, or ``broad'' $P1$ (2.2) from Berenson et al.). For
instance, in the example below, $T_3$ should never see $a=1$, and, if
$T_2$ aborts, $T_3$ should never read $a=3$:
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)~w_x(2)
\\T_2 &: w_x(3)\\
T_3 &: r_x(a)\\[-2em]
\end{align*}
It is fairly easy for a HAT system to prevent ``Dirty Reads'': if each
transaction never writes uncommitted data to the database, then
transactions will never read each others' dirty data. As a simple
solution, clients can buffer their writes until they commit, or,
alternatively, can send them to servers, who will not deliver their
value to other readers until notified that the writes have been
committed. This implementation does not provide recency or
monotonicity guarantees but satisfies the implementation-agnostic
definition.

@@

\end{tikzpicture}
\label{fig:hat-order}
\caption{HAT, sticky HAT (in boxes), and unavailable models (circled)
  from Table~\protect\ref{table:hatcompared} compared
  graphically. Directed edges represent ordering by model
  strength. Models that do not share a common ancestor can be
  simultaneously achieved, and the resulting availability is that of
  the weakest available model in the combination.}
\label{fig:hatcompared}
\end{figure}

@@

Composition cost also varies by combination. For instance, Charron-Bost
 proved that, to capture causality between $N$ communicating
processes, standard vector-based approaches face an upper bound of
$O(N)$ storage per write~\cite{charron-bost}. This means that, with
$100K$ clients, each write might be accompanied by $100K$ timestamps
per vector. This is difficult to scale. By compromising on
availability (e.g., treating a datacenter as a linearizable cluster),
this overhead can be reduced~\cite{eiger}, but it is much
cheaper to provide, say, read your writes, than full causal
consistency.

>>>>

\textbf{Read Committed} isolation is particularly important in
practice as it is the default of many DBMSs. Implementations differ,
with some based on long-duration exclusive locks and short-duration
read locks~\cite{gray-isolation} and others based on multiple
versions. The implementations typically have recency and monotonicity
properties beyond the simple meaning of the name, which is what is
expressed in the implementation-agnostic definition: under Read
Committed, transactions should do not access uncommitted or
intermediate versions of data items. This prohibits both ``Dirty
Writes'', as above, and also ``Dirty Reads'' phenomena.  This
isolation is Adya's PL-2, and formalized by prohibiting Adya's $G1a-c$
(or ANSI's $P1$, or ``broad'' $P1$ (2.2) from Berenson et al.). For
instance, in the example below, $T_3$ should never see $a=1$, and, if
$T_2$ aborts, $T_3$ should never read $a=3$:
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)~w_x(2)
\\T_2 &: w_x(3)\\
T_3 &: r_x(a)\\[-2em]
\end{align*}
It is fairly easy for a HAT system to prevent ``Dirty Reads'': if each
transaction never writes uncommitted data to the database, then
transactions will never read each others' dirty data. As a simple
solution, clients can buffer their writes until they commit, or,
alternatively, can send them to servers, who will not deliver their
value to other readers until notified that the writes have been
committed. This implementation does not provide recency or
monotonicity guarantees but satisfies the implementation-agnostic
definition.

@@

\end{tikzpicture}
\label{fig:hat-order}
\caption{HAT, sticky HAT (in boxes), and unavailable models (circled)
  from Table~\protect\ref{table:hatcompared} compared
  graphically. Directed edges represent ordering by model
  strength. Models that do not share a common ancestor can be
  simultaneously achieved, and the resulting availability is that of
  the weakest available model in the combination.}
\label{fig:hatcompared}
\end{figure}\vspace{-1em}

@@

Composition cost may also vary by combination. For instance, Charron-Bost
 proved that, to capture causality between $N$ communicating
processes, standard vector-based approaches face an upper bound of
$O(N)$ storage per write~\cite{charron-bost}. This means that, with
$100K$ clients, each write might be accompanied by $100K$ timestamps
per vector. This is difficult to scale. By compromising on
availability (e.g., treating a datacenter as a linearizable cluster),
this overhead can be reduced~\cite{eiger}, but it is much
cheaper to provide, say, read your writes, than full causal
consistency.
